A high-repetition-rate, fast temperature-programmed gas chromatograph and its on-line coupling to a supercritical fluid chromatograph (SFC×GC)

### 2019-09-14 document.tex
Paper as submitted to RSI

\title{A high-repetition-rate, fast temperature programmed gas chromatograph and
its application in comprehensive \texorpdfstring{(supercritical fluid × gas)}{supercritical fluid x gas}
chromatography } %Title of paper

@@

\keywords{high speed temperature programmed GC, subambient GC, fast process GC}
\begin{abstract}

We present a fast gas chromatographic system that can be used as a second
dimension in comprehensive (supercritical fluid × gas) chromatography
(SFC×GC). The temperature of the short (1 metre long) capillary column is
controlled by a resistively heated coaxial stainless steel tube. The electrical
resistance and therefore temperature of the stainless steel tube is measured by
continuous monitoring of the current/voltage ratio. Highly repeatable heating
rates of up to 2100 C\textdegree/min (35 C\textdegree/s) are obtained, which
should be sufficient for the most demanding fast chromatograms. To reduce the
cooling time between temperature programs the column is cooled by injecting
evaporating carbon dioxide into the space between the coaxial heater and the
column. This gives cooling rates of 5100 C\textdegree/min (85 C\textdegree/s)
which allows quick succession of temperature programs. More repeatable heating
profiles with stable GC retention times together with faster cooling are
significant improvements on previous SFC×GC systems. Cycle times of four
gas chromatograms per minute could readily be achieved, which allows efficient
coupling to high-resolution stop-flow SFC in the first dimension.

We demonstrate the fast chromatograph by separating fatty acid methyl esters by
SFC×GC chromatography, yielding fatty acid profile information which
would be useful in the food and biodiesel industries.

@@

In comprehensive 2D chromatography the second dimension chromatography must be
\textit{fast}. The theory and practice of improving the speed of GC is well
developed\cite*{Cramers1999,Korytar2002}. Typically, short columns and high
carrier flows rates are used, which reduce retention times by reducing the
void time. The use of narrow-bore columns and thin films reduce run times by
improving mass transfer between the mobile and stationary phases, allowing
higher linear flow rates of the carrier gas.

@@

In this work we used a coaxial heater, in the form of a thin-walled stainless
steel tube that surrounds the capillary column.  This tube is resistively heated
by controlled direct current. Due to the low thermal mass the heating is rapid,
and since the resistance of the tube indicates its temperature there is no need
for external temperature measuring devices. The benefit of this design is that
it allows the use of commercially-available fused silica columns with their wide
range of internal diameters and selectivities. Columns can be exchanged without
altering the temperature feedback system and therefore the rapid heating
characteristics. Additionally, the co-axial space between the capillary column
and the stainless steel tube can be filled with coolant to rapidly cool the
column.

@@

Agilent\texttrademark{} markets a `low thermal mass' column, which includes
collinear heating wire and a collinear sensing element bundled with a short
silica column. These units cool down faster than normal air-baths, but only to
ambient temperature. For short columns a cooling time of \SI{1.3}{\minute} has
been reported\cite{Luong2006}. These columns are available with only a limited
range of stationary phases.

The Zip Scientific\texttrademark{} solution is forced convection. Their
`GC-Chaser' system\cite{ZipScientific} uses a blast of (chilled) room air to the
GC oven to cool it down after each GC run, allowing the use of conventional
fused silica columns. In an example\cite{ZipScientific2008} it reduces cooling
time from 16 minutes to 7 minutes. Although this can significantly decrease the
cost per analysis in a high-throughput laboratory, the cooling is not fast
enough for 2D chromatography.

@@

Cryogenic coolers for GC ovens are available, but are intended to cool the oven
to low temperatures for the analysis of volatile compounds and not necessarily
to improve cycle times. In previous SFC×GC work in our
laboratories\cite{Venter2004, Venter2006} the cryogenic function of a Varian
3300 gas chromatograph was used to cool down the column oven at the beginning of
each run, using large quantities of coolant in the process. The resistively
heated column cooling required 30 seconds to get back to the starting
temperature in this cool environment. We estimate that a single SFC×GC
run consumed an unacceptable 15 kg of carbon dioxide coolant.

@@

After the sample is injected, the SFC runs for an optional uninterrupted period,
usually briefer than the void time. No data is collected during this time. Then
the stop valve closes and the cooling system activates to cool the GC column.
Once the column is at the set temperature, the stop valve opens and the SFC
eluate exits the restrictor into the hot splitless inlet of the GC, where it
evaporates and is transported on to the column. The vapor-phase material
is trapped in the cold stationary phase. Once the desired amount
of the SFC eluate has been collected, the stop valve closes. Some time is
allowed to pass so that the last of the evaporated SFC eluate flushes into the
column. Then the split vent of the GC inlet opens for a short period, which
vents excess carbon dioxide to the atmosphere. This normalizes the pressure in
the inlet (having been raised  by the high flow rate of CO$_2$ from the SFC) and
allows the carbon dioxide to be replaced by the pressure-controlled hydrogen
carrier gas. Once the split vent closes the GC run starts, which of course means
the start of the temperature program, which desorbs the analyte trapped on the
head of the cold column. At the end of the GC run, the cooling process starts,
cooling the column in a second or two. The next cycle then starts, marked by the
opening of SFC the stop valve.

@@

The inlet was an unmodified Varian\texttrademark{} 1075 split/splitless inlet.
The temperature control of the inlet was by the original electronics, controlled
from the Varian 3300 front panel. The split vent valve was disconnected from the
gas chromatograph, and controlled from the control computer described below.

@@

The coaxial heater was made of a 940 mm length of 304 stainless steel, obtained
from MIFAM (Milanówek, Poland). It had an outside diameter of 1.06 mm and an
inside diameter of 0.80 mm.

@@

The T-piece blocks were machined from brass. The stainless steel tube of the
coaxial heater was brazed onto the block. The cooling CO$_2$ was let in from one
side, and on the opposite side a thermocouple probe inserted into a blind hole
measured the temperature of the block. A micro-union using a metal ferrule
(Restek\texttrademark{} SilTite $\mu$-Union Connectors), brazed to the block,
was used to seal the capillary column. An electrical contact was added, to which
an electrical conductor could be soft-soldered. Four holes through the length of
the block allowed the fitting of four 220V, 100 W heater cartridges (Hotset),
providing each of the two blocks 400W of heating.

\begin{figure}
	\includegraphics[width=8.5cm]{./Figures/Cutaway_with_symbols.png}
	\caption["The T-piece block"]{A cutaway diagram of the inlet/outlet outlet T-piece blocks, showing the (a) cryogenic coolant inlet/outlet, (b) the coaxial heater, (c) the capillary column, (d) the mico-union seal, (e) the cartridge heater socket, (f) the thermocouple socket, and (g) the electrical connection.}%
	\label{fig:Manifold}
\end{figure}

The chromatographic column entered the T-piece block via the bottom port with the
coaxial heater tube, and passed out through the top port and into the inlet or
detector.

@@

In the left-hand side of the T-piece block was the inlet or outlet port for the
carbon dioxide coolant.

@@

The cryogenic carbon dioxide (Afrox, Tec (Wet)) was supplied from a cylinder
equipped with a dip tube. It flowed to a coil heat exchanger bathed in a cold
(-5 \textdegree C) propylene glycol heat transfer fluid, which was mounted on
top of the Varian GC. From the heat exchanger the carbon dioxide flowed down to
the cryo shut-off valve (ASCO, RedHat brand). This arrangement ensured a supply
of liquid CO$_2$ at the shut-off valve, without which some gaseous CO$_2$ might
have to vent through the coaxial heater before the liquid CO$_2$ coolant could
reach it. Downstream of the shut-of valve there was a 10-turn metering valve,
which controlled the flow of carbon dioxide into the coaxial heater for
evaporative cooling. At the outlet end of the coaxial heater the carbon dioxide
escaped through the T-piece block into the atmosphere.

@@

A controlled direct current was used to heat the coaxial heater. The heating
circuit consisted of the coaxial heater in series with a reference resistor and
a ballast resistor. The reference resistor was constructed out of parallel
stainless steel wires with an open construction to promote heat dissipation with
the aim of keeping its temperature constant. The ballast resistor was of stainless
steel wire in a bifilar winding with an air core, to minimize electromagnetic
interference and encourage heat dissipation.

@@

\begin{figure}
\includegraphics[width=8.5cm]{./Figures/ResistanceMeasurement.pdf}%
\caption{\label{Circuit}A simplified circuit diagram of the resistance measuring circuit.}%
\end{figure}

@@

A two-stage control system controlled the current. The first stage of the
control system was a PID controller implemented in LabVIEW. The process variable
was the temperature of the coaxial heater, as calculated from its electrical
resistance, and the set value was determined by the desired chromatographic
temperature program. The manipulated variable was the voltage of the
digital-to-analog converter converter (DAC) of the data acquisition board. The
final stage was an electronic system, where the voltage applied to the coaxial
heater was the process variable, and the set value was a voltage provided by the
DAC. The manipulated variable was the base current of the power transistors.
In summary, the software set a voltage, which set the power of the coaxial
heater. The two-stage design allows for switching of the power to the coaxial
heater from computer control to manual control by a potentiometer on the
instrument front panel. Being able to control the power of the coaxial heater
manually is extremely useful during development and fault-finding.

@@

\begin{figure}
\includegraphics[width=8.5cm]{./Figures/1-16.jpg}%
\caption{\label{fig:ThermImg}A thermal image of the coaxial heater.}%
\end{figure}

@@

\begin{figure}
	\includegraphics[width=8.5cm]{./Figures/cp.pdf}%
	\caption{\label{fig:Repeatability}A graph indicating that temperature gradients are modest, stable and repeatable}%
\end{figure}

@@

While it is conceptually possible to have a fast temperature program by using
the voltage/current ratio alone as a temperature indicator and controlled
variable, it would be hard to design chromatographic methods or translate
methods without a knowledge of the real temperature. Therefore we went to some
effort to calibrate the coaxial heater.

To calibrate a thermometer, the thermometer under test and its calibration
environment must be in equilibrium. This requirement is hard to meet in our
calibration, because a single point (the junction of the thermocouple) needs to
be in equilibrium with the length of the coaxial heater at all times during fast
heating. This equilibrium would be very difficult to maintain, especially
because the uniformity of heating cannot be guaranteed, and is additionally
influenced by the presence of the T-piece heaters at the high temperature of the
GC inlet and detector --- for a totally uniform heating of the coaxial heater the
T-piece blocks would have to be at the same temperature as the coaxial heater.

@@

For reasons not understood, the calibration curve is oddly curved, and we
arbitrarily fitted a B-spline to the data points. This procedure can be
automated\cite{WenniZheng2012}. Points from this B-spline curve were then
extracted to create a calibration curve, from which an estimate of the
temperature can be calculated by linear interpolation\cite{Possolo2017}.

\begin{figure}
\includegraphics[width=8.5cm]{./Figures/2015_04_24_Calibration.pdf}%
\caption{\label{TempCal}A calibration curve.}%
\end{figure}

@@

To achieve good resolution in temperature-programmed chromatography, a good rule
of thumb for the heating rate is \SI{10}{\celsius} per void
time\cite{Blumberg2000}. With hydrogen as carrier gas in a
\SI{0.25}{\milli\metre} i.d. column of the optimum linear velocity is
\SI{40}{\centi\metre\per\second}. Then the void time of a \SI{1}{\metre} column
is \SI{2.5}{\s}, so the recommended heating rate is \SI{10}{\celsius} per
\SI{2.5}{\s}, which is \SI{4}{\celsius}/s, or \SI{240}{\celsius\per\minute}.
Since in fast chromatography excess peak resolution is traded for shorter run
times, higher than optimum flow rates will be used, so higher heating rates will
be required. For narrower columns the flow will be even higher so heating rates
of more than 1000 C\textdegree/min will be required.

The attainable heating rate is a function of power to the heater and of the
temperature required. (In the instrument presented here power was limited by the
voltage output range of the DAC of the data acquisition board, and not by the
power of the heater.) We were able to maintain controlled heating rates of 
2500 C\textdegree /min up to 400 \textdegree C, or up to 4000 C\textdegree /min
up to 350 \textdegree C. (Figure \ref{fig:MaxHeatingRate})

The heating rates attainable seem to be sufficient for the most demanding fast
gas chromatography.

\begin{figure}
\includegraphics[width=8.5cm]{./Figures/high_rate_heating.pdf}%

\caption{\label{fig:MaxHeatingRate}A graph of 9 identical, consecutive temperature ramps overlaid. 
The heating rate is 4000 C\textdegree /min. The temperature follows the setpoint faithfully up to \SI{350}{\celsius}}

@@

Using evaporating carbon dioxide as a powerful coolant, it was possible to cool
down the column at a rate of 5100 C$^\circ$/min. The cooling rate depends on the
setting of the metering valve, with the lowest coaxial heater temperatures and
smallest gradients achieved at an intermediate flow of about
\SI{30}{\gram\per\minute}.

@@

Results obtained from creating a 2D blank chromatogram with two alkanes in the
mobile phase is summarized in Table \ref{tab:table1}. Peak widths were about
\SI{500}{\milli\second}. The relative standard deviations (RSD) of retention
times were similar to those obtained in GC×GC\cite{Shellie2002}.

@@

\begin{ruledtabular}
\begin{tabular}{lllll}
Compound & n & t\textsubscript{r} (s) & S.D. of t\textsubscript{r} (s)& R.S.D. of t\textsubscript{r} (\%)\\
\hline
Dodecane & 73 & 5.069 & 0.0231 & 0.456\\
Hexadecane & 73 & 6.584 & 0.0516 & 0.784\\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{figure}
\includegraphics[width=8.5cm]{./Figures/Interpretation.png}%

@@

Prof. Walter Meyer of the Department of Physics at the University of Pretoria
supplied the idea for the electronics that measure the resistance of the
stainless steel heating tube. Nico van Vuuren machined the T-piece blocks and
their mounting rails. We thank Restek for the generous donation of the packed
silica columns. Reinhardt Heymans from FLIR Systems recorded the thermal images
at no cost. David Masemula kept the laboratory supplied with consumables.

>>>>

\title{A high-repetition-rate, fast temperature-programmed gas chromatograph and
its on-line coupling to a supercritical fluid chromatograph (SFC×GC)}
%Title of paper

@@

\keywords{high speed temperature programmed GC, sub-ambient GC, fast process GC}
\begin{abstract}

We present a fast gas chromatographic system that can be used as a second
dimension in comprehensive (supercritical fluid × gas) chromatography (SFC×GC).
The temperature of the short (1 metre long) capillary column is controlled by a
resistively heated coaxial stainless-steel tube. The electrical resistance and
therefore temperature of the stainless-steel tube is measured by continuous
monitoring of the current/voltage ratio. Highly repeatable heating rates of up
to \SI{2100}{\celsius\per\minute} (\SI{35}{\celsius\per\second}) are obtained,
which should be high enough for the most demanding fast chromatograms. To reduce
the cooling time between temperature programs the column is cooled by injecting
evaporating carbon dioxide into the space between the coaxial heater and the
column. This gives cooling rates of \SI{5100}{\celsius\per\minute}
(\SI{85}{\celsius\per\minute}) which allows quick succession of temperature
programs. More repeatable heating profiles with stable GC retention times
together with faster cooling are significant improvements on previous SFC×GC
systems. Cycle times of four gas chromatograms per minute could readily be
achieved, which allows efficient coupling to high-resolution stop-flow SFC in
the first dimension. We demonstrate the fast chromatograph by separating fatty
information which would be useful in the food and biodiesel industries.

@@

In comprehensive 2D chromatography the second-dimension chromatography must be
\textit{fast}. The theory and practice of improving the speed of GC is well
developed\cite*{Cramers1999,Korytar2002}. Typically, short columns and high
carrier flows rates are used, which reduce retention times by reducing the
void time. The use of narrow-bore columns and thin films reduce run times by
improving mass transfer between the mobile and stationary phases, allowing
higher linear flow rates of the carrier gas.

@@

In this work we used a coaxial heater, in the form of a thin-walled
stainless-steel tube that surrounds the capillary column.  This tube is
resistively heated by controlled direct current. Due to the low thermal mass the
heating is rapid, and since the resistance of the tube indicates its temperature
there is no need for external temperature measuring devices. The benefit of this
design is that it allows the use of commercially-available fused silica columns
with their wide range of internal diameters and selectivities. Columns can be
exchanged without altering the temperature feedback system and therefore the
rapid heating characteristics. Additionally, the co-axial space between the
capillary column and the stainless steel tube can be filled with coolant to
rapidly cool the column.

@@

Agilent\texttrademark{} markets a `low thermal mass' column, which includes a
collinear heating wire and a collinear sensing element bundled with a short
silica column. These units cool down faster than normal air-baths, but only to
ambient temperature. For short columns a cooling time of \SI{1.3}{\minute} has
been reported\cite{Luong2006}. These columns are available with only a limited
range of stationary phases.

The Zip Scientific\texttrademark{} solution is forced convection. Their
`GC-Chaser' system uses a blast of (chilled) room air to the
GC oven to cool it down after each GC run, allowing the use of conventional
fused silica columns. In an example\cite{ZipScientific2008} it reduces cooling
time from 16 minutes to 7 minutes. Although this can significantly decrease the
cost per analysis in a high-throughput laboratory, the cooling is not fast
enough for 2D chromatography.

@@

Cryogenic coolers for GC ovens are available but are intended to cool the oven
to low temperatures for the analysis of volatile compounds and not necessarily
to improve cycle times. In previous SFC×GC work in our
laboratories\cite{Venter2004, Venter2006} the cryogenic function of a Varian
3300 gas chromatograph was used to cool down the column oven at the beginning of
each run, using large quantities of coolant in the process. The resistively
heated column cooling required 30 seconds to get back to the starting
temperature in this cool environment. We estimate that a single SFC×GC run
consumed an unacceptable 15 kg of carbon dioxide coolant.

@@

After the sample is injected, the SFC runs for an optional uninterrupted period,
usually briefer than the void time. No data is collected during this time. Then
the stop valve closes, and the cooling system activates to cool the GC column.
Once the column is at the set temperature, the stop valve opens and the SFC
eluate exits the restrictor into the hot splitless inlet of the GC, where it
evaporates and is transported on to the column. The vapor-phase material
is trapped in the cold stationary phase. Once the desired amount
of the SFC eluate has been collected, the stop valve closes. Some time is
allowed to pass so that the last of the evaporated SFC eluate flushes into the
column. Then the split vent of the GC inlet opens for a short period, which
vents excess carbon dioxide to the atmosphere. This normalizes the pressure in
the inlet (having been raised  by the high flow rate of CO$_2$ from the SFC) and
allows the carbon dioxide to be replaced by the pressure-controlled hydrogen
carrier gas. Once the split vent closes the GC run starts, which of course means
the start of the temperature program, which desorbs the analyte trapped on the
head of the cold column. At the end of the GC run, the cooling process starts,
cooling the column in a second or two. The next cycle then starts, marked by the
opening of SFC the stop valve.

@@

The inlet was an unmodified Varian\texttrademark{} 1075 split/splitless inlet.
The temperature control of the inlet was by the original electronics, controlled
from the Varian 3300 front panel. The split vent valve was disconnected from the
gas chromatograph and controlled from the control computer described below.

@@

The coaxial heater was made of a 940 mm length of stainless steel (SAE 304
grade), obtained from MIFAM (Milanówek, Poland). It had an outside diameter of
1.06 mm and an inside diameter of 0.80 mm.

@@

The T-piece blocks were machined from brass. The stainless-steel tube of the
coaxial heater was brazed onto the block. The cooling CO$_2$ was let in from one
side, and on the opposite side a thermocouple probe inserted into a blind hole
measured the temperature of the block. A micro-union using a metal ferrule
(Restek\texttrademark{} SilTite $\mu$-Union Connectors), brazed to the block,
was used to seal the capillary column. An electrical contact was added, to which
an electrical conductor could be soft-soldered. Four holes through the length of
the block allowed the fitting of four 220V, 100 W heater cartridges (Hotset),
providing each of the two blocks 400W of heating.

\begin{figure}
	\includegraphics[width=\textwidth]{./Figures/Cutaway_with_symbols.png}
	\caption["The T-piece block"]{A cutaway diagram of the inlet/outlet outlet T-piece blocks, showing the (a) cryogenic coolant inlet/outlet, (b) the coaxial heater, (c) the capillary column, (d) the mico-union seal, (e) the cartridge heater socket, (f) the thermocouple socket, and (g) the electrical connection.}%
	\label{fig:Manifold}
\end{figure}

The chromatographic column entered the T-piece block via the bottom port with the
coaxial heater tube and passed out through the top port and into the inlet or
detector. In the left-hand side of the T-piece block was the inlet or outlet port for the
carbon dioxide coolant.

@@

@@

The cryogenic carbon dioxide (Afrox, Tec (Wet)) was supplied from a cylinder
equipped with a dip tube. It flowed to a coil heat exchanger bathed in a cold
\SI{-5}{\celsius}) propylene glycol heat transfer fluid, which was mounted on
top of the Varian GC. From the heat exchanger the carbon dioxide flowed down to
the cryo shut-off valve (ASCO, RedHat brand). This arrangement ensured a supply
of liquid CO$_2$ at the shut-off valve, without which some gaseous CO$_2$ might
have to vent through the coaxial heater before the liquid CO$_2$ coolant could
reach it. Downstream of the shut-of valve there was a 10-turn metering valve,
which controlled the flow of carbon dioxide into the coaxial heater for
evaporative cooling. At the outlet end of the coaxial heater the carbon dioxide
escaped through the T-piece block into the atmosphere.

@@

A controlled direct current was used to heat the coaxial heater. The heating
circuit consisted of the coaxial heater in series with a reference resistor and
a ballast resistor. The reference resistor was constructed out of parallel
stainless-steel wires with an open construction to promote heat dissipation with
the aim of keeping its temperature constant. The ballast resistor was of
stainless-steel wire in a bifilar winding with an air core, to minimize
electromagnetic interference and encourage heat dissipation.

@@

\begin{figure}
\includegraphics[width=\textwidth]{./Figures/ResistanceMeasurement.pdf}%
\caption{\label{Circuit}A simplified circuit diagram of the resistance measuring circuit.}%
\end{figure}

@@

A two-stage control system controlled the current. The first stage of the
control system was a PID controller implemented in LabVIEW. The process variable
was the temperature of the coaxial heater, as calculated from its electrical
resistance, and the set value was determined by the desired chromatographic
temperature program. The manipulated variable was the voltage of the
digital-to-analog converter converter (DAC) of the data acquisition board. The
final stage was an electronic system, where the voltage applied to the coaxial
heater was the process variable, and the set value was a voltage provided by the
DAC. The manipulated variable was the base current of the power transistors.
In summary, the software set a voltage, which set the power of the coaxial
heater. The two-stage design allows for switching of the control of the coaxial
heater from computer control to manual control by a potentiometer on the
instrument front panel. Being able to control the power of the coaxial heater
manually is extremely useful during development and fault-finding.

@@

\begin{figure}
\includegraphics[width=\textwidth]{./Figures/1-16.jpg}%
\caption{\label{fig:ThermImg}A thermal image of the coaxial heater.}%
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\textwidth]{./Figures/cp.pdf}%
	\caption{\label{fig:Repeatability}A graph indicating that temperature gradients are modest, stable and repeatable}%
\end{figure}

@@

While it is conceptually possible to have a fast temperature program by using
the voltage/current ratio alone as a temperature indicator and controlled
variable, it would be hard to design chromatographic methods or translate
methods without a knowledge of the real temperature. Therefore, we went to some
effort to calibrate the coaxial heater.

To calibrate a thermometer, the thermometer under test and its calibration
environment must be in equilibrium. This requirement is hard to meet in our
calibration, because a single point (the junction of the thermocouple) needs to
be in equilibrium with the length of the coaxial heater during the whole of fast
heating cycle. This equilibrium would be very difficult to maintain, especially
because the uniformity of heating cannot be guaranteed, and is additionally
influenced by the presence of the T-piece heaters at the high temperature of the
GC inlet and detector --- for a totally uniform heating of the coaxial heater
the T-piece blocks would have to be at the same temperature as the coaxial
heater. Acknowledging these limitations, we performed a calibration procedure
assuming that the thermocouple junction was in equilibrium with the coaxial
heater, which was assumed to be of uniform temperature over its length.

@@

For reasons not understood, the calibration (see Figure \ref{TempCal}) curve is
oddly curved, and we arbitrarily fitted a B-spline to the data points. This
procedure can be automated\cite{WenniZheng2012}. Points from this B-spline curve
were then extracted to create a calibration curve, from which an estimate of the
temperature can be calculated by linear interpolation\cite{Possolo2017}.

\begin{figure}
\includegraphics[width=\textwidth]{./Figures/2019_07_12_B-spline_fit_CPlot.pdf}%

\caption{\label{TempCal}A calibration curve of the coaxial heater. The markers
were obtained from a B-spline fitted manually to the curve. The temperature is
obtained by linear interpolation between markers.}%

@@

To achieve good resolution in temperature-programmed chromatography, a good rule
of thumb for the heating rate is \SI{10}{\celsius} per void
time\cite{Blumberg2000}. With hydrogen as carrier gas in a
\SI{0.25}{\milli\metre} i.d. column of the optimum linear velocity is
\SI{40}{\centi\metre\per\second}. Then the void time of a \SI{1}{\metre} column
is \SI{2.5}{\s}, so the recommended heating rate is \SI{10}{\celsius} per
\SI{2.5}{\s}, which is \SI{4}{\celsius}/s, or \SI{240}{\celsius\per\minute}.
Since in fast chromatography excess peak resolution is traded for shorter run
times, higher than optimum flow rates will be used, so higher heating rates will
be required. For narrower columns the flow will be even higher so heating rates
of more than \SI{1000}{\celsius\per\minute} will be required.

The attainable heating rate is a function of power to the heater and of the
temperature required. (In the instrument presented here power was limited by the
voltage output range of the DAC of the data acquisition board, and not by the
power of the heater.) We were able to maintain controlled heating rates of
\SI{2500}{\celsius\per\minute} up to \SI{400}{\celsius}, or up to \SI{4000}{\celsius\per\minute}
up to \SI{350}{\celsius}. (Figure \ref{fig:MaxHeatingRate}.) The heating rates
attainable seem to be high enough for the most demanding fast gas
chromatography.

\begin{figure}
\includegraphics[width=\textwidth]{./Figures/high_rate_heating.pdf}%

\caption{\label{fig:MaxHeatingRate}A graph of 9 identical, consecutive
temperature ramps overlaid. The heating rate is \SI{4000}{\celsius\per\minute}.
The temperature follows the setpoint faithfully up to \SI{350}{\celsius}}

@@

Using evaporating carbon dioxide as a powerful coolant, it was possible to cool
down the column at a rate of \SI{5100}{\celsius\per\minute}. The cooling rate
depends on the setting of the metering valve, with the lowest coaxial heater
temperatures and smallest gradients achieved at an intermediate flow of about
\SI{30}{\gram\per\minute}.

@@

Results obtained from creating a 2D blank chromatogram with two alkanes in the
mobile phase is summarized in Table \ref{tab:table1}. The peak widths were about
\SI{500}{\milli\second}, so the \SI{20}{\milli\second} standard deviation for
hexadecane means that the variation in retention time is only about
\SI{10}{\percent} of the peak width. The relative standard deviations (RSD) of
retention times were similar to those obtained in GC×GC\cite{Shellie2002}.

@@

\begin{ruledtabular}
\begin{tabular}{lllll}
Compound & n & t\textsubscript{r} (s) & S.D. of t\textsubscript{r} (s)& R.S.D. of t\textsubscript{r} (\%)\\
\hline
Dodecane & 73 & 5.07 & 0.023 & 0.46\\
Hexadecane & 73 & 6.58 & 0.052 & 0.78\\
\end{tabular}
\end{ruledtabular}
\end{table}

@@

\begin{figure}
\includegraphics[width=\textwidth]{./Figures/Interpretation.png}%

@@

Prof. Walter Meyer of the Department of Physics at the University of Pretoria
supplied the idea for the resistance-measuring electronic circuit. Nico van
Vuuren machined the T-piece blocks and their mounting rails. We thank Restek for
the generous donation of the packed silica columns. Reinhardt Heymans from FLIR
Systems recorded the thermal images at no cost. David Masemula kept the
laboratory supplied with consumables.
